---
title: "note.How to use the WDI library in R for World Bank data"
author: "JI"
date: "6/1/2021"
output:
  html_document:
   toc: true
   number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd(getwd())
library(WDI)
library(ggplot2)
library(plotly)
library(dplyr)
```

<style type="text/css">
h1 {color:#1e1a51}
h2 {color:#ffbd4a}
h3 {color:#1a2451}
</style>

## Introduction 

The WorldBank(R) provides open data for hundreds of indicators.  These indicators are available for most countries spanning many years.  You can learn more at https://datacatalog.worldbank.org/ 

The `WDI` library provides easy access to the data. 

## Looking up an indicator 

The World Bank data that is available through this API provides many variable including _iso2c_ country codes, names of the variables, and many more.  The `WDISearch()` command allows strings to be searched for in these variables.  In the code block below a search for _maternal mortality ratio_ is made in the _name_ variable.

```{r looking for indicator}
WDIsearch(string = "maternal mortality ratio", 
          field = "name",
          short = FALSE)
```

## Looking for the data

You could get more information via `?WDI`
```{r searching for the data}
df <- data.frame(
  WDI(country= c("US","BR","ZA"), 
    indicator="SH.STA.MMRT",
    start=2005, end=2015,extra = FALSE))
head(df)
```

the full list of variables is shown below 
```{r}
names(WDI(country = "US",
          indicator = "SH.STA.MMRT",
          start = 1990,
          end = 2020, 
          extra = TRUE))
```

## Filtering data 
we could use `filter()` to subset data from the dataset  

```{r}
brazil <- df %>% filter(iso2c == "BR")
usa <- df %>% filter(iso2c == "US")
za <- df %>% filter(iso2c == "ZA")
```

## Plotting the data 

the code block creates three traces for a single plot using `plotly()`

```{r}
trace0 <- brazil$SH.STA.MMRT
trace1 <- usa$SH.STA.MMRT
trace2 <- za$SH.STA.MMRT
dts <- brazil$year
df.plot <- data.frame(dts,trace0,trace1, trace2)

p1 <- plot_ly(df.plot,
              x = ~dts,
              y = ~trace0,
              name = "Brazil",
              type = "scatter",
              mode = "lines+markers") %>% 
  add_trace(y = ~trace1,
            name = "USA",
            type = "scatter",
            mode = "lines+markers") %>% 
  add_trace(y = ~trace2,
            name = "RSA",
            type = "scatter",
            mode = "lines+markers")%>% 
  layout(title = "Maternal mortality per 100,000 births",
         xaxis = list(title = "Year",
                      zeroline = FALSE),
         yaxis = list(title = "Count",
                      zeroline = FALSE))
p1
```

## Using ggplot to plot the data

the idea is to comparing if `ggplot2` could also plot the data as `plotly()`

```{r ggplot2 packages}
plot2 <- df %>% 
  ggplot(aes(x = year, y = SH.STA.MMRT, color = iso2c)) +
  geom_line()+
  geom_point()+
  scale_x_discrete(limits = c(2005:2015))+# change the x_scale 
  ggtitle("Maternal mortality per 100,000 births")+
  xlab("Year")+
  ylab("Count")
  
plot2
```

## Conclusion

The WDI library is extremely easy to use, yet powerful enough to bring a massive open data resource to you desktop.  Learn more about it at: https://cran.r-project.org/web/packages/WDI/WDI.pdf 

If you would like to know more about maternal mortality rates in South Africa, you can view this issue of the South African Medical Journal: http://www.samj.org.za/index.php/samj/issue/view/215/showToc 


